---
title: "Final_Exam"
author: "Vy Nguyen"
date: "4/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(TSA)
library(tseries)
```

## 1
a. For ARIMA(0,1,1), we should perform ARIMA(1,1,1) and ARIMA(0,1,2)
b. 
c. IMA(1,1) with $\theta$ = 0.6
d. $\hat{Y}_t(2) = \mu + \phi^2 (Y_t - \mu)$  
$\hat{Y}_t(2) = 0 + 0.6^2(Y_{100} - 0)$  
$\hat{Y}_t(2) = 0.72$
e. False????????
f. ADF p-value 0.7 means do not reject null hypothesis, therefore we take difference. the ACF for the series has a very slow decay, and the PACF for the difference series has significant spikes at lags 1 and 2 means that AR(2) model would be most consistent with the findings.
Therefore, we choose ARIMA(2,1,0)
g. True

## 2
a. The  fitted model :
$logY_t = logY_{t-1} + e_t + 0.2956e_{t-1}$
b.  ML estimate $\hat{\theta}$ = -0.2956 is significantly different
from 0 since the approximate 95 percent confidence interval is (-0.2956−1.96∗0.0693, -0.2956+1.96∗0.0693) = (−0.431428, −0.159772).
c. The figure on the RIGHT may correspond to MMSE forecast and 95%
prediction intervals according to your model in (a).
Since we fit IMA(1,1) model to the data, we assume that the original data is non-stationary, and as expected from a nonstationary process, the estimated standard errors (also on the log scale) increase
as l increases.
d. MMSE forecast on the original scale is $e^{5+1/2*0.2^2}$ = 151.411
The 95 percent confidence interval of the log-scale is (4.608,5.392)
Therefore 95 percent confidence interval of the original forecast:
($e^{4.608}$,$e^{5.392}$) = (100.283,219.642)

## 3
```{r}
ts2022 = as.ts(read.table(file = "ts2022.txt"))
```

```{r}
ts.plot(ts2022, type="o", main="TS 2022", ylab="Unit", xlab="Time")
```
```{r}
acf(ts2022,main="Sample ACF")
```
```{r}
adf.test(ts2022)
```
a. The process in the plot does not assemble a stationary process. Also, the sample ACF decays slowly every 5 lags, which also indicates that the process is not stationary.The p-value of the ADF test is greater than 0.05, so we fail to reject the null hypothesis, which states that the process is not stationary.
b. 
```{r}
par(mfrow=c(1,2))
plot(diff(ts2022),ylab="First Difference of Gasoline Prices",xlab="Time",type="o")
acf(diff(ts2022),main="Sample ACF")
```

```{r}
adf.test(diff(ts2022))
```
For the first differences of ts2022 data, the time series plot looks like a mean zero white noise process. The p-value of the ADF test is smaller than 0.01, so we reject the null hypothesis and we claim that there is sufficient evidence to support that the process is stationary.
```{r}
pacf(diff(ts2022),main="Sample PACF")
```
```{r}
eacf(diff(ts2022))
```
The sample ACF suggests that we should use MA(3)
The EACF suggest we should use ARMA(0,3) or ARMA(2,1)
Therefore, ARIMA(0,1,3) or ARIMA(0,1,1) will be a candidate model for the dataset.
c.
We will fit ARIMA(0,1,3)
```{r}
arima(ts2022,order=c(2,1,1),method="ML")
```

The 95% CI for $\hat{\theta}_1$ = (−0.146,0.057) which is not significantly different from zero, therefore we have to try another model.
We will fit ARIMA(0,1,1)

```{r}
arima(ts2022,order=c(2,1,1),method="ML")
```




